---
title: "Autumn activity hotspots"
format: html
editor: visual
---

```{r setup }
#| echo: false
#| warning: false

library(tidyverse)
library(ggplot2)
library(dplyr)
library(suncalc)
#library(suntools)
library(lubridate)
library(leaflet.minicharts)
library(leaflet)
library(leaflet.extras)

# read data ---------------------------------------------------------------

cm <- read.csv("cm_2025.csv")
overview_summary <- read.csv("overview_2025.csv")

# cut data from July

cm$DATE <- as.Date(cm$DATE)
cm$DATE_12 <- as.Date(cm$DATE_12)

cm <- cm %>%
  filter(DATE >= "2025-07-01")

# add enil calls to overview and change pnat calls to only autumn too, add different timeframes for calls in the overview table.

overview_summary <- cm %>%
  group_by(Site) %>%
  summarise(
    # EPTNIL
    enilaut   = sum(autoid == "EPTNIL"),
    enil_aug   = sum(autoid == "EPTNIL" & DATE >= "2025-08-15" & DATE <= "2025-08-31"),
    enil_sep1  = sum(autoid == "EPTNIL" & DATE >= "2025-09-01" & DATE < "2025-09-15"),
    enil_sep2  = sum(autoid == "EPTNIL" & DATE >= "2025-09-16" & DATE <= "2025-09-30"),
    enil_oct = sum(autoid == "EPTNIL" & DATE >= "2025-10-01"),
    
    # PIPNAT
    pnataut   = sum(autoid == "PIPNAT"),
    pnat_aug   = sum(autoid == "PIPNAT" & DATE >= "2025-08-15" & DATE <= "2025-08-31"),
    pnat_sep1  = sum(autoid == "PIPNAT" & DATE >= "2025-09-01" & DATE < "2025-09-15"),
    pnat_sep2  = sum(autoid == "PIPNAT" & DATE >= "2025-09-16" & DATE <= "2025-09-30"),
    pnat_oct = sum(autoid == "PIPNAT" & DATE >= "2025-10-01")
  ) %>%
  left_join(overview_summary, ., by = "Site")

```


## Autumn activity

Trying to make a visual picture based on coastal monitoring data from 2025. 
Activity for 3 species- northern bat, parti-coloured bat and nathusius pipistrelle, on 4 different time periods- II half of August, I half of September, II half of September and October,

## Locations

Coastal monitoring sites with a bigger emphasis on sites we could mist net in.
Not a lot of activity north of Stavanger. Interesting areas: Stavanger area lakes, coast in Jæren (around Orre and Brusand), Bryne, Brusand, Lista. 

## Northern bat
### Overall activity

```{r enil}
#| echo: false
#| label: fig-enil
#| warning: false
#| fig-cap: Overall northern bat activity


enil <- cm %>%
  filter(autoid=="EPTNIL")

ggplot(enil %>% 
         droplevels(), aes(x= DATE_12, color = taxa)) + 
  stat_count( geom = "point", size= 4, alpha = 0.90, color = "royalblue") +
  ylab("5s Recordings per night") + xlab("Month") +
  ggtitle("Autumn activity 2025") +
  scale_x_date(date_breaks = "1 month", date_labels = "%b") +
  theme_minimal()

```

### Activity at different times
#### End of August

Still active, most active sites were...

```{r enil map august}
#| label: fig-enil-map-a
#| echo: false
#| warning: false
#| fig-cap: Second part of August northern bat activity

tilesURL <- "http://server.arcgisonline.com/ArcGIS/rest/services/Canvas/World_Light_Gray_Base/MapServer/tile/{z}/{y}/{x}"

basemap <- leaflet(width = "100%", height = "400px") %>%
  addTiles(tilesURL)

basemap %>%
  addMinicharts(
    overview_summary$longitude, overview_summary$latitude,
    chartdata = overview_summary$enil_aug, 
    showLabels = TRUE,
    width = 45
  )

```


#### Beginning of September

Not that active anymore, but we could still find them at...

```{r enil map sept1}
#| label: fig-enil-map-s1
#| echo: false
#| warning: false
#| fig-cap: First part of September northern bat activity

tilesURL <- "http://server.arcgisonline.com/ArcGIS/rest/services/Canvas/World_Light_Gray_Base/MapServer/tile/{z}/{y}/{x}"

basemap <- leaflet(width = "100%", height = "400px") %>%
  addTiles(tilesURL)

basemap %>%
  addMinicharts(
    overview_summary$longitude, overview_summary$latitude,
    chartdata = overview_summary$enil_sep1, 
    showLabels = TRUE,
    width = 45
  )

```

#### End of September, October

```{r enil map sept2}
#| label: fig-enil-map-s2
#| echo: false
#| warning: false
#| fig-cap: Second part of September northern bat activity

tilesURL <- "http://server.arcgisonline.com/ArcGIS/rest/services/Canvas/World_Light_Gray_Base/MapServer/tile/{z}/{y}/{x}"

basemap <- leaflet(width = "100%", height = "400px") %>%
  addTiles(tilesURL)

basemap %>%
  addMinicharts(
    overview_summary$longitude, overview_summary$latitude,
    chartdata = overview_summary$enil_sep2, 
    showLabels = TRUE,
    width = 45
  )

```

```{r enil map oct}
#| label: fig-enil-map-o
#| echo: false
#| warning: false
#| fig-cap: October northern bat activity

tilesURL <- "http://server.arcgisonline.com/ArcGIS/rest/services/Canvas/World_Light_Gray_Base/MapServer/tile/{z}/{y}/{x}"

basemap <- leaflet(width = "100%", height = "400px") %>%
  addTiles(tilesURL)

basemap %>%
  addMinicharts(
    overview_summary$longitude, overview_summary$latitude,
    chartdata = overview_summary$enil_oct, 
    showLabels = TRUE,
    width = 45
  )

```

## Parti-coloured bat
### Overall activity

```{r vmur}
#| echo: false
#| label: fig-vmur
#| warning: false
#| fig-cap: Overall parti-coloured bat activity


vmur <- cm %>%
  filter(autoid=="VESMUR")

ggplot(vmur %>% 
         droplevels(), aes(x= DATE_12, color = taxa)) + 
  stat_count( geom = "point", size= 4, alpha = 0.90, color = "royalblue") +
  ylab("5s Recordings per night") + xlab("Month") +
  ggtitle("Autumn activity 2025") +
  scale_x_date(date_breaks = "1 month", date_labels = "%b") +
  theme_minimal()

```

Sites that the species is active around: 


## Nathusius pipistrelle
### Overall activity

```{r pnat}
#| echo: false
#| label: fig-pnat
#| warning: false
#| fig-cap: Overall nathusius pipistrelle activity


pnat <- cm %>%
  filter(autoid=="PIPNAT")

ggplot(pnat %>% 
         droplevels(), aes(x= DATE_12, color = taxa)) + 
  stat_count( geom = "point", size= 4, alpha = 0.90, color = "royalblue") +
  ylab("5s Recordings per night") + xlab("Month") +
  ggtitle("Autumn activity 2025") +
  scale_x_date(date_breaks = "1 month", date_labels = "%b") +
  theme_minimal()

```

### Activity at different times
#### End of August

```{r pnat map august}
#| label: fig-pnat-map-a
#| echo: false
#| warning: false
#| fig-cap: Second part of August nathusius pipistrelle activity

tilesURL <- "http://server.arcgisonline.com/ArcGIS/rest/services/Canvas/World_Light_Gray_Base/MapServer/tile/{z}/{y}/{x}"

basemap <- leaflet(width = "100%", height = "400px") %>%
  addTiles(tilesURL)

basemap %>%
  addMinicharts(
    overview_summary$longitude, overview_summary$latitude,
    chartdata = overview_summary$pnat_aug, 
    showLabels = TRUE,
    width = 45
  )

```

Not so active yet, some bats at...

#### Beginning of September

Very active time, especially... What stands out: 

```{r pnat map sept1}
#| label: fig-pnat-map-s1
#| echo: false
#| warning: false
#| fig-cap: First part of September nathusius pipistrelle activity

tilesURL <- "http://server.arcgisonline.com/ArcGIS/rest/services/Canvas/World_Light_Gray_Base/MapServer/tile/{z}/{y}/{x}"

basemap <- leaflet(width = "100%", height = "400px") %>%
  addTiles(tilesURL)

basemap %>%
  addMinicharts(
    overview_summary$longitude, overview_summary$latitude,
    chartdata = overview_summary$pnat_sep1, 
    showLabels = TRUE,
    width = 45
  )

```

#### End of September

Activity shift more to ...

```{r pnat map sept2}
#| label: fig-pnat-map-s2
#| echo: false
#| warning: false
#| fig-cap: Second part of September nathusius pipistrelle activity

tilesURL <- "http://server.arcgisonline.com/ArcGIS/rest/services/Canvas/World_Light_Gray_Base/MapServer/tile/{z}/{y}/{x}"

basemap <- leaflet(width = "100%", height = "400px") %>%
  addTiles(tilesURL)

basemap %>%
  addMinicharts(
    overview_summary$longitude, overview_summary$latitude,
    chartdata = overview_summary$pnat_sep2, 
    showLabels = TRUE,
    width = 45
  )

```

#### October

```{r pnat map o}
#| label: fig-pnat-map-o
#| echo: false
#| warning: false
#| fig-cap: October nathusius pipistrelle activity

tilesURL <- "http://server.arcgisonline.com/ArcGIS/rest/services/Canvas/World_Light_Gray_Base/MapServer/tile/{z}/{y}/{x}"

basemap <- leaflet(width = "100%", height = "400px") %>%
  addTiles(tilesURL)

basemap %>%
  addMinicharts(
    overview_summary$longitude, overview_summary$latitude,
    chartdata = overview_summary$pnat_oct, 
    showLabels = TRUE,
    width = 45
  )

#another

pal <- colorNumeric(
  palette = "YlOrRd",
  domain = overview_summary$pnat_oct
)

leaflet(overview_summary) %>%
  addTiles(tilesURL) %>%
  addCircleMarkers(
    lng = ~longitude,
    lat = ~latitude,
    radius = ~sqrt(pnat_oct)/2,
    color = ~pal(pnat_oct),
    fillOpacity = 0.7,
    stroke = FALSE,
    popup = ~paste0(
      "<b>Site:</b> ", Site,
      "<br><b>PIPNAT:</b> ", pnat_oct
    )
  ) %>%
  addLegend(
    "bottomright",
    pal = pal,
    values = ~pnat_oct,
    title = "PIPNAT calls"
  )



```


```{r change}

overview_summary <- overview_summary %>%
  mutate(
    change = pnat_sep2 - pnat_aug
  )

pal <- colorNumeric(
  palette = "RdYlGn",
  domain = overview_summary$change
)

leaflet(overview_summary) %>%
  addTiles(tilesURL) %>%
  addCircleMarkers(
    lng = ~longitude,
    lat = ~latitude,
    radius = ~sqrt(pnat_sep2) / 2,
    color = ~pal(change),
    fillColor = ~pal(change),
    fillOpacity = 0.8,
    stroke = FALSE,
    popup = ~paste0(
      "<b>Site:</b> ", Site,
      "<br><b>Aug:</b> ", pnat_aug,
      "<br><b>Sep 1–15:</b> ", pnat_sep1,
      "<br><b>Sep 16–30:</b> ", pnat_sep2,
      "<br><b>Change:</b> ", change
    )
  ) %>%
  addLegend(
    "bottomright",
    pal = pal,
    values = ~change,
    title = "Change in PIPNAT"
  )

```

