---
title: "Autumn activity hotspots"
format: html
theme: pulse
editor: visual
---

```{r setup }
#| echo: false
#| warning: false

library(tidyverse)
library(ggplot2)
library(dplyr)
library(suncalc)
#library(suntools)
library(lubridate)
library(leaflet.minicharts)
library(leaflet)
library(leaflet.extras)

# read data ---------------------------------------------------------------

cm <- read.csv("cm_2025.csv")
overview_summary <- read.csv("overview_2025.csv")

# cut data from July

cm$DATE <- as.Date(cm$DATE)
cm$DATE_12 <- as.Date(cm$DATE_12)

cm <- cm %>%
  filter(DATE >= "2025-07-01")

# add enil calls to overview and change pnat calls to only autumn too, add different timeframes for calls in the overview table.

overview_summary <- cm %>%
  group_by(Site) %>%
  summarise(
    # EPTNIL
    enilaut   = sum(autoid == "EPTNIL"),
    enil_aug   = sum(autoid == "EPTNIL" & DATE >= "2025-08-15" & DATE <= "2025-08-31"),
    enil_sep1  = sum(autoid == "EPTNIL" & DATE >= "2025-09-01" & DATE < "2025-09-15"),
    enil_sep2  = sum(autoid == "EPTNIL" & DATE >= "2025-09-16" & DATE <= "2025-09-30"),
    enil_oct = sum(autoid == "EPTNIL" & DATE >= "2025-10-01"),
    
    # PIPNAT
    pnataut   = sum(autoid == "PIPNAT"),
    pnat_aug   = sum(autoid == "PIPNAT" & DATE >= "2025-08-15" & DATE <= "2025-08-31"),
    pnat_sep1  = sum(autoid == "PIPNAT" & DATE >= "2025-09-01" & DATE < "2025-09-15"),
    pnat_sep2  = sum(autoid == "PIPNAT" & DATE >= "2025-09-16" & DATE <= "2025-09-30"),
    pnat_oct = sum(autoid == "PIPNAT" & DATE >= "2025-10-01"),
    
    #VESMUR
    vmursept = sum (autoid == "VESMUR" & DATE >= "2025-09-01")
    
  ) %>%
  left_join(overview_summary, ., by = "Site")

```


Trying to make a visual picture based on coastal monitoring data from 2025. 
Activity for two species- northern bat and nathusius pipistrelle, on 4 different time periods- II half of August, I half of September, II half of September and October,

## Locations

Coastal monitoring sites with a bigger emphasis on sites we could mist net in.
Not a lot of activity north of Stavanger. Interesting areas: Stavanger area lakes, coast in Jæren (around Orre and Brusand), Bryne, Lista. 

:::{style="background-color: #d5d6ea; padding: 20px; border-radius: 5px;"}

## Northern bat
### Overall activity

```{r enil}
#| echo: false
#| label: fig-enil
#| warning: false
#| fig-cap: Overall northern bat activity


enil <- cm %>%
  filter(autoid=="EPTNIL")

ggplot(enil %>% 
         droplevels(), aes(x= DATE_12, color = taxa)) + 
  stat_count( geom = "point", size= 4, alpha = 0.90, color = "royalblue") +
  ylab("5s Recordings per night") + xlab("Month") +
  ggtitle("Autumn activity 2025") +
  scale_x_date(date_breaks = "1 month", date_labels = "%b") +
  theme_minimal()

```

### Activity at different times
#### End of August

Still active, most active sites were the lakesites- Mosvatnet, Stokkalandsvatnet, Taksdalsvatnet and Smokkavatnet. Also Bedehus stands out as an active site for August.

```{r enil map august}
#| label: fig-enil-map-a
#| echo: false
#| warning: false
#| fig-cap: Second part of August northern bat activity

tilesURL <- "http://server.arcgisonline.com/ArcGIS/rest/services/Canvas/World_Light_Gray_Base/MapServer/tile/{z}/{y}/{x}"

basemap <- leaflet(width = "100%", height = "494px") %>%
  addTiles(tilesURL)

basemap %>%
  addMinicharts(
    overview_summary$longitude, overview_summary$latitude,
    chartdata = overview_summary$enil_aug, 
    showLabels = TRUE,
    width = 45
  )

```

#### Beginning of September

Not that active anymore, only comparable locations to August are Store Stokkavatnet and Mosvatnet. 

```{r enil map sept1}
#| label: fig-enil-map-s1
#| echo: false
#| warning: false
#| fig-cap: First part of September northern bat activity

basemap %>%
  addMinicharts(
    overview_summary$longitude, overview_summary$latitude,
    chartdata = overview_summary$enil_sep1, 
    showLabels = TRUE,
    width = 45
  )

```

#### Change from August to September

The activity of northern bat decreases in all locations from August to September, the size of the circles represents the activity left of the second part of September.

```{r enil map change}
#| label: fig-enil-map-c
#| echo: false
#| warning: false
#| fig-cap: Change of northern bat activity from end of August to September

overview_summary <- overview_summary %>%
  mutate(
    change1 = enil_sep2 - enil_aug
  )

pal <- colorNumeric(
  palette = "RdYlBu",
  domain = overview_summary$change1
)

leaflet(overview_summary) %>%
  addTiles(tilesURL) %>%
  addCircleMarkers(
    lng = ~longitude,
    lat = ~latitude,
    radius = ~sqrt(enil_sep2) / 2,
    color = ~pal(change1),
    fillColor = ~pal(change1),
    fillOpacity = 0.8,
    stroke = FALSE,
    popup = ~paste0(
      "<b>Site:</b> ", Site,
      "<br><b>Aug:</b> ", enil_aug,
      "<br><b>Sep 1–15:</b> ", enil_sep1,
      "<br><b>Sep 16–30:</b> ", enil_sep2,
      "<br><b>Change:</b> ", change1
    )
  ) %>%
  addLegend(
    "bottomright",
    pal = pal,
    values = ~change1,
    title = "Change in northern bat activity"
  )

```

```{r enil map oct}
#| label: fig-enil-map-o
#| echo: false
#| output: false
#| warning: false
#| fig-cap: October northern bat activity

tilesURL <- "http://server.arcgisonline.com/ArcGIS/rest/services/Canvas/World_Light_Gray_Base/MapServer/tile/{z}/{y}/{x}"

basemap <- leaflet(width = "100%", height = "400px") %>%
  addTiles(tilesURL)

basemap %>%
  addMinicharts(
    overview_summary$longitude, overview_summary$latitude,
    chartdata = overview_summary$enil_oct, 
    showLabels = TRUE,
    width = 45
  )

```
:::
<!-- ## Parti-coloured bat -->
<!-- ### Overall activity -->

<!-- ```{r vmur} -->
<!-- #| echo: false -->
<!-- #| label: fig-vmur -->
<!-- #| warning: false -->
<!-- #| fig-cap: Overall parti-coloured bat activity -->


<!-- vmur <- cm %>% -->
<!--   filter(autoid=="VESMUR") -->

<!-- ggplot(vmur %>%  -->
<!--          droplevels(), aes(x= DATE_12, color = taxa)) +  -->
<!--   stat_count( geom = "point", size= 4, alpha = 0.90, color = "royalblue") + -->
<!--   ylab("5s Recordings per night") + xlab("Month") + -->
<!--   ggtitle("Autumn activity 2025") + -->
<!--   scale_x_date(date_breaks = "1 month", date_labels = "%b") + -->
<!--   theme_minimal() -->

<!-- ``` -->

<!-- Sites that the species is active around in September: Smokkavatnet, Taksdalsvatnet and Bryne. -->

::: {style="background-color: #c2e5d3; padding: 20px; border-radius: 5px;"}
## Nathusius pipistrelle

### Overall activity

```{r pnat}
#| echo: false
#| label: fig-pnat
#| warning: false
#| fig-cap: Overall nathusius pipistrelle activity


pnat <- cm %>%
  filter(autoid=="PIPNAT")

ggplot(pnat %>% 
         droplevels(), aes(x= DATE_12, color = taxa)) + 
  stat_count( geom = "point", size= 4, alpha = 0.90, color = "royalblue") +
  ylab("5s Recordings per night") + xlab("Month") +
  ggtitle("Autumn activity 2025") +
  scale_x_date(date_breaks = "1 month", date_labels = "%b") +
  theme_minimal()

```


### Activity at different times

#### Change from August to September

```{r pnat map august}
#| label: fig-pnat-map-a
#| output: false
#| echo: false
#| warning: false
#| fig-cap: Second part of August nathusius pipistrelle activity


basemap %>%
  addMinicharts(
    overview_summary$longitude, overview_summary$latitude,
    chartdata = overview_summary$pnat_aug, 
    showLabels = TRUE,
    width = 45
  )

```

```{r pnat change aug-sept}
#| label: fig-pnat-map-c1
#| echo: false
#| warning: false
#| fig-cap: Change of troll activity from end of August to September

overview_summary <- overview_summary %>%
  mutate(
    change2 = pnat_sep1 - pnat_aug
  )

pal <- colorNumeric(
  palette = "RdYlGn",
  domain = overview_summary$change2
)

leaflet(overview_summary) %>%
  addTiles(tilesURL) %>%
  addCircleMarkers(
    lng = ~longitude,
    lat = ~latitude,
    radius = ~sqrt(pnat_sep1) / 2,
    color = ~pal(change2),
    fillColor = ~pal(change2),
    fillOpacity = 0.8,
    stroke = FALSE,
    popup = ~paste0(
      "<b>Site:</b> ", Site,
      "<br><b>Aug:</b> ", pnat_aug,
      "<br><b>Sep 1–15:</b> ", pnat_sep1,
      "<br><b>Change:</b> ", change2
    )
  ) %>%
  addLegend(
    "bottomright",
    pal = pal,
    values = ~change2,
    title = "Change in troll activity"
  )

```

In almost all locations, the activity rose, active locations in the beginning of September: Steinodden, Brusand, Lista, Store Stokkavatnet.The radius of the point shows the comparative activity in September


#### Change in September

In some locations, orange tones show that the activity does not change, but it does increase in Store Stokkavatnet.The radius of the point shows the comparative activity in the end of September.

```{r pnat change sept}
#| label: fig-pnat-map-c2
#| echo: false
#| warning: false
#| fig-cap: Change of troll activity in September

overview_summary <- overview_summary %>%
  mutate(
    change3 = pnat_sep2 - pnat_sep1
  )

pal <- colorNumeric(
  palette = "RdYlGn",
  domain = overview_summary$change3
)

leaflet(overview_summary) %>%
  addTiles(tilesURL) %>%
  addCircleMarkers(
    lng = ~longitude,
    lat = ~latitude,
    radius = ~sqrt(pnat_sep2) / 2,
    color = ~pal(change3),
    fillColor = ~pal(change3),
    fillOpacity = 0.8,
    stroke = FALSE,
    popup = ~paste0(
      "<b>Site:</b> ", Site,
      "<br><b>Sep 1-15:</b> ", pnat_sep1,
      "<br><b>Sep 16–30:</b> ", pnat_sep2,
      "<br><b>Change:</b> ", change3
    )
  ) %>%
  addLegend(
    "bottomright",
    pal = pal,
    values = ~change3,
    title = "Change in troll activity"
  )

```

```{r pnat map sept1}
#| label: fig-pnat-map-s1
#| echo: false
#| output: false
#| warning: false
#| fig-cap: First part of September nathusius pipistrelle activity

basemap %>%
  addMinicharts(
    overview_summary$longitude, overview_summary$latitude,
    chartdata = overview_summary$pnat_sep1, 
    showLabels = TRUE,
    width = 45
  )

```

```{r pnat map sept2}
#| label: fig-pnat-map-s2
#| echo: false
#| output: false
#| warning: false
#| fig-cap: Second part of September nathusius pipistrelle activity

tilesURL <- "http://server.arcgisonline.com/ArcGIS/rest/services/Canvas/World_Light_Gray_Base/MapServer/tile/{z}/{y}/{x}"

basemap <- leaflet(width = "100%", height = "400px") %>%
  addTiles(tilesURL)

basemap %>%
  addMinicharts(
    overview_summary$longitude, overview_summary$latitude,
    chartdata = overview_summary$pnat_sep2, 
    showLabels = TRUE,
    width = 45
  )

```

#### October

In warm October, there is still some activity left in the south and around Stavanger.

```{r pnat map o}
#| label: fig-pnat-map-o
#| echo: false
#| warning: false
#| fig-cap: October nathusius pipistrelle activity


basemap %>%
  addMinicharts(
    overview_summary$longitude, overview_summary$latitude,
    chartdata = overview_summary$pnat_oct, 
    showLabels = TRUE,
    width = 45
  )

```

#### Time of night

Activity pattern during September nights in 7 potential mist netting locations. Watch out: overall activity can be different because of some outlier nights.

Mosvatnet (CM-06) - Activity drops on cold nights and rainy weather.
Lista (CM-11) - Some very active nights around 9th and 10th of September, rest of the month not so active.
Steinodden (CM-12) - Some active night around 12, 14, 15 September. 
Fuglevika (CM-13) - Active night 9th of Sept.
Store Stokkavatnet (CM-17) - A lot more activity in the last week of the month.
Brusand (CM-25) - More activity 17th to 23rd of September.
Bryne (CM-56) - More stable activity, but overall lowel levels than others. 

```{r pnat diurnal }
#| label: fig-pnat-diu
#| echo: false
#| warning: false
#| fig-cap: Activity timing in September

pnat %>%
  filter(
    DATE_12 >= "2025-09-01" & DATE_12 <= "2025-09-30",
    Site %in% c("CM-06", "CM-17", "CM-56", "CM-25", "CM-11", "CM-12", "CM-13")
  ) %>%
    mutate(
    HOUR_plot = ifelse(HOUR < 14, HOUR + 24, HOUR)
  ) %>%
  ggplot() +
  aes(x = HOUR_plot) +
  geom_histogram(
    binwidth = 1,
    boundary = 0,
    fill = "#112446"
  ) +
  scale_x_continuous(
    breaks = c(14, 16, 18, 20, 22, 24, 26, 28, 30, 32, 34, 36),
    labels = c(14, 16, 18, 20, 22, 24, 2, 4, 6, 8, 10, 12)
  ) +
  ylab("5s Recordings") + xlab("Hour") +
  ggtitle("Pnat activity times in september") +
  theme_minimal() +
  facet_wrap(vars(Site))


```
```{r pnat diurnal site}
#| label: fig-pnat-diu
#| echo: false
#| output: false
#| warning: false
#| fig-cap: Activity timing in September per day

# change site to the one you want to view

pnat %>%
  filter(
    DATE_12 >= "2025-09-01" & DATE_12 <= "2025-09-30",
    Site %in% c("CM-56")
  ) %>%
    mutate(
    HOUR_plot = ifelse(HOUR < 14, HOUR + 24, HOUR)
  ) %>%
  ggplot() +
  aes(x = HOUR_plot) +
  geom_histogram(
    binwidth = 1,
    boundary = 0,
    fill = "#112446"
  ) +
  scale_x_continuous(
    breaks = c(14, 18, 22, 26, 30, 34),
    labels = c(14, 18, 22, 2, 6, 10)
  ) +
  ylab("5s Recordings") + xlab("Hour") +
  ggtitle("Pnat activity times in september") +
  theme_minimal() +
  facet_wrap(vars(DATE))


```

:::
