---
title: "Overview of data collected in the field season of 2025 for coastal monitoring"
author: "Maris Pärn"
format:
  html:
    embed-resources: true
editor: visual
---

```{r setup}
#| echo: false
#| warning: false

#setup

library(tidyverse)
library(ggplot2)
library(dplyr)
library(lubridate)
library(MetBrewer)
library(hms)
library(leaflet.minicharts)
library(leaflet)
library(tidyr)



# read necessary data in

overview_summary <- read.csv("overview_2025.csv")
cm <- read.csv("cm_2025.csv") 
missing_dates <- read.csv("missing_dates_2025.csv")
nodata_2025 <- read.csv("Missing_nights_nodata.csv")

# overview figures

available_dates <- cm %>%
  mutate(ActualDate = as.Date(DATE_12)) %>%
  reframe(ActualDate = unique(ActualDate), .by = "Site")

complete_dates <- overview_summary %>% 
  mutate(
    BeginDate = as.Date(date_deployed),
    EndDate   = coalesce(as.Date(date_retrieved), as.Date("2025-10-22")) - 1,
    full_dates = map2(BeginDate, EndDate, ~ seq.Date(from = .x, to = .y, by = "day"))
  ) %>%
  select(Site, full_dates) %>%
  unnest(full_dates) %>%
  rename(ExpectedDate = full_dates)

```

## Overview

###Year in numbers

Number of detectors: 50\

Number of detectors that survived the season: 49\

Reasons of detector failure: seawater damage, storms, microphone losing sensitivity, teft.\

Date of first deployment: `r min(overview_summary$date_deployed)`\

Date of last deployment: `r max(overview_summary$date_deployed)`\

Date of first retrieval: `r min(overview_summary$date_retrieved)`\

Date of last retrieval: `r max(overview_summary$date_retrieved)`\

Days detectors were on the field: `r nrow(available_dates)`\

Days with missing data: `r nrow(missing_dates)`\

\% of missing data from all data: ca `r  100 * nrow(missing_dates)/ nrow(complete_dates)`\

Reasons of missing data:\

1\) No bats, no noise detected (later Nodata): 134\

2\) SD card full: 244\

3\) Batteries empty: 17\

4\) Detector dead: 113\

5\) Detector stolen: 55\

6\) SD card not formatted: 34\

Average time detector deployed: `r mean(overview_summary$days_active)`\

```{r preparing missing dates data}
#| echo: false
#| warning: false

complete_dates <- overview_summary |>  
  mutate(
    BeginDate = as.Date(date_deployed),
    EndDate   = as.Date(date_retrieved) - 1,
    full_dates = map2(BeginDate, EndDate, ~ seq.Date(from = .x, to = .y, by = "day"))
  ) |> 
  select(Site, full_dates) |> 
  unnest(full_dates) |> 
  rename(ExpectedDate = full_dates)

available_dates <- cm %>%
  mutate(ActualDate = as.Date(DATE_12)) %>%
  reframe(ActualDate = unique(ActualDate), .by = "Site")

available_dates <- available_dates %>%
  mutate(Type = "Observed")

missing_dates$X <- NULL

missing_dates <- missing_dates %>%
  rename(
    Date = ExpectedDate
  )

available_dates <- available_dates %>%
  rename(
    Date = ActualDate
  )

missing_dates <- missing_dates %>%
  mutate(Type = "Missing") %>% 
  mutate(Date=ymd(Date))


all_dates <- bind_rows(available_dates, missing_dates)

100 * nrow(missing_dates)/ nrow(complete_dates)
```

```{r data preparation for plots in this document}
#| echo: false
#| warning: false

# missing dates plot
  # add type NoData (aka no recordings) to Observed and Missing

nodata_2025 <- nodata_2025 %>%
  mutate(ExpectedDate = as.Date(ExpectedDate, format = "%d/%m/%Y")) %>%
  select(Site, ExpectedDate, NoData) %>%
  rename(
    Date = ExpectedDate,
    Type = NoData
  )

nodata_2025 <- nodata_2025[-which(nodata_2025$Type == ""), ]

all_dates <- all_dates |>
  left_join(nodata_2025, by = c("Site", "Date"), suffix = c("", "_new")) |>
  mutate(Type = if_else(!is.na(Type_new), Type_new, Type)) |>
  select(-Type_new)

# noise plot

noise_coast <- cm[,c(13,17)] #autoid + site
noise_coast$type <- overview_summary$type[match(noise_coast$Site, overview_summary$Site)]

# autoid plot

auto_id_data <- cm[,c(10,13, 15, 17)] #date12, autoid, match ratio and site
auto_id_data$type <- overview_summary$type[match(auto_id_data$Site, overview_summary$Site)]

# combined bat calls plot

cm$DATE <- as.Date(cm$DATE)
cm$DATE_12 <- as.Date(cm$DATE_12)

# activity plot
  # adding a column "type"

cm$type <- overview_summary$type[match(cm$Site, overview_summary$Site)]


```

```{r pnat data preparation}
#| echo: false
#| warning: false

# choose PIPNAT

cm_pnat <- cm %>%
  filter(autoid == "PIPNAT") 

cm_pnat_c <- cm_pnat |>
  filter(type == "coast") %>%
  filter(month(DATE_12) %in% c(5, 6, 7, 8, 9, 10))

# add median dates

median_dates_all <- cm_pnat %>%
  mutate(DATE_12 = as.Date(DATE_12)) %>%
  group_by(Site) %>%
  summarise(median = median(DATE_12, na.rm = TRUE))

median_dates_all$latitude <-overview_summary$latitude[match(median_dates_all$Site, overview_summary$Site)]

median_dates_all <- median_dates_all %>%
  arrange(latitude) %>%
  mutate(Site = factor(Site, levels = Site))

cm_pnat_c <- cm_pnat_c %>%
  mutate(TIME=as_hms(TIME))

cm_pnat_c <- cm_pnat_c %>%
  mutate(TIME_12=as_hms(TIME_12))


# median(cm_pnat_c$DATE.12, na.rm = FALSE)

median_dates <- cm_pnat_c %>%
  mutate(DATE_12 = as.Date(DATE_12)) %>%
  group_by(Site) %>%
  summarise(median = median(DATE_12, na.rm = TRUE))

median_dates$latitude <-overview_summary$latitude[match(median_dates$Site, overview_summary$Site)]

median_dates <- median_dates %>%
  arrange(latitude) %>%
  mutate(Site = factor(Site, levels = Site))

cm_pnat_c <- cm_pnat_c %>%
  mutate(TIME=as_hms(TIME))

cm_pnat_c <- cm_pnat_c %>%
  mutate(TIME_12=as_hms(TIME_12))

# half a month

cm_pnat_c <- cm_pnat_c %>%
  mutate(
    HalfMonth = case_when(
      lubridate::day(DATE) <= 15 ~
        paste0(lubridate::year(DATE), "-", sprintf("%02d", lubridate::month(DATE)), "-H1"),
      TRUE ~
        paste0(lubridate::year(DATE), "-", sprintf("%02d", lubridate::month(DATE)), "-H2")
    )
  )

#check for duplicates
  #dim(cm_pnat_c[duplicated(cm_pnat_c$filename),])[1]

# cm to date



```

## Deployment times

This is a graph that shows the days with missing data and how long the detectors were in the field.

```{r missing dates plot}
#| echo: false
#| label: fig-missing
#| warning: false
#| fig-cap: Missing dates 2024

ggplot(all_dates, aes(x = Date, y = Site, color = Type)) +
  geom_point(size = 2) +
  scale_color_manual(values = c("Observed" = "azure3", "NoData" = "darkgrey", "Missing" = "red")) +
  theme_minimal() +
  scale_y_discrete(expand = expansion(mult = c(0.01, 0.01)))+
  labs(title = "Observed and Missing Dates for 2025")


```

### Nathusius' pipistrelle activity
ggplot(all_dates, aes(x = Date, y = Site, color = Type)) +
  geom_point(size = 2) +
  scale_color_manual(values = c("Observed" = "azure3", "NoData" = "darkgrey", "Missing" = "red")) +
  theme_minimal() +
  scale_y_discrete(expand = expansion(mult = c(0.01, 0.01)))+
  labs(title = "Observed and Missing Dates for 2025")

We can see that in our sites, the activity peaks in September.

```{r pnat}
#| echo: false
#| label: fig-pnat
#| warning: false
#| fig-cap: Pnat calls

ggplot(cm %>% 
    filter(autoid == "PIPNAT") %>% 
    droplevels(), aes(x= DATE_12, color = taxa)) + 
  stat_count( geom = "point", size= 4, alpha = 0.90, color = "royalblue") +
    ylab("Recordings per night") + xlab("Month") +
   ggtitle("Overall amount of calls 2025") +
    scale_x_date(date_breaks = "1 month", date_labels = "%b") +
    theme_minimal()
```

```{r pnat 3}
#| echo: false
#| label: fig-pnat-3
#| warning: false
#| fig-cap: Comparison of Pnat activity in different sites
#| fig-subcap:
#| - "Activity on coast"
#| - "Activity around lakes"
#| - "Activity inland"
#| layout-ncol: 3
#| column: page


ggplot(cm %>% 
    filter(autoid == "PIPNAT",
            type == "coast") %>%
    droplevels(), aes(x= DATE_12, color = taxa)) + 
  stat_count( geom = "point", size= 4, alpha = 0.90, color = "royalblue") +
    ylab("Recordings per night") + xlab("Month") +
   ggtitle("Overall activity on the coast") +
    scale_x_date(date_breaks = "1 month", date_labels = "%b") +
    theme_minimal()

ggplot(cm %>% 
    filter(autoid == "PIPNAT",
            type == "lake") %>%
    droplevels(), aes(x= DATE_12, color = taxa)) + 
  stat_count( geom = "point", size= 4, alpha = 0.90, color = "royalblue") +
    ylab("Recordings per night") + xlab("Month") +
   ggtitle("Overall activity around lakes") +
    scale_x_date(date_breaks = "1 month", date_labels = "%b") +
    theme_minimal()

ggplot(cm %>% 
    filter(autoid == "PIPNAT",
            type == "inland") %>%
    droplevels(), aes(x= DATE_12, color = taxa)) + 
  stat_count( geom = "point", size= 4, alpha = 0.90, color = "royalblue") +
    ylab("Recordings per night") + xlab("Month") +
   ggtitle("Overall activity inland") +
    scale_x_date(date_breaks = "1 month", date_labels = "%b") +
    theme_minimal()
```

### Soprano pipistrelle activity

Activity is higher leading up to the September peak, but we can still clearly see a peak that occurs earlier than that of Nathusius’ pipistrelles.

```{r ppyg}
#| echo: false
#| label: fig-ppyg
#| warning: false
#| fig-cap: Ppyg calls


ggplot(cm %>% 
    filter(autoid == "PIPPYG") %>% 
    droplevels(), aes(x= DATE_12, color = taxa)) + 
  stat_count( geom = "point", size= 4, alpha = 0.90, color = "pink4") +
    ylab("Recordings per night") + xlab("Month") +
   ggtitle("Overall activity 2025") +
    scale_x_date(date_breaks = "1 month", date_labels = "%b") +
    theme_minimal()
```

```{r ppyg 3}
#| echo: false
#| label: fig-ppyg-3
#| warning: false
#| fig-cap: Comparison of Ppyg activity in different sites
#| fig-subcap:
#| - "Activity on coast"
#| - "Activity around lakes"
#| - "Activity inland"
#| layout-ncol: 3
#| column: page

ggplot(cm %>% 
    filter(autoid == "PIPPYG",
            type == "coast") %>%
    droplevels(), aes(x= DATE_12, color = taxa)) + 
  stat_count( geom = "point", size= 4, alpha = 0.90, color = "pink4") +
    ylab("Recordings per night") + xlab("Month") +
   ggtitle("Overall activity on the coast") +
    scale_x_date(date_breaks = "1 month", date_labels = "%b") +
    theme_minimal()

ggplot(cm %>% 
    filter(autoid == "PIPPYG",
            type == "lake") %>%
    droplevels(), aes(x= DATE_12, color = taxa)) + 
  stat_count( geom = "point", size= 4, alpha = 0.90, color = "pink4") +
    ylab("Recordings per night") + xlab("Month") +
   ggtitle("Overall activity around lakes") +
    scale_x_date(date_breaks = "1 month", date_labels = "%b") +
    theme_minimal()

ggplot(cm %>% 
    filter(autoid == "PIPPYG",
            type == "inland") %>%
    droplevels(), aes(x= DATE_12, color = taxa)) + 
  stat_count( geom = "point", size= 4, alpha = 0.90, color = "pink4") +
    ylab("Recordings per night") + xlab("Month") +
   ggtitle("Overall activity inland") +
    scale_x_date(date_breaks = "1 month", date_labels = "%b") +
    theme_minimal()
```

### Daubentons' bat activity

```{r mdau}
#| echo: false
#| label: fig-mdau
#| warning: false
#| fig-cap: Mdau calls

ggplot(cm %>% 
    filter(autoid == "MYODAU") %>% 
    droplevels(), aes(x= DATE_12, color = taxa)) + 
  stat_count( geom = "point", size= 4, alpha = 0.90, color = "skyblue") +
    ylab("Recordings per night") + xlab("Month") +
   ggtitle("Overall activity") +
    scale_x_date(date_breaks = "1 month", date_labels = "%b") +
    theme_minimal()
```

### Northern bat activity

We can see a high level of avtivity throughout the recording period up until the beginning of september, when activity suddenly drops.

```{r enil}
#| echo: false
#| label: fig-enil
#| warning: false
#| fig-cap: AUTO-ID Enil calls

ggplot(cm %>% 
    filter(autoid == "EPTNIL") %>% 
    droplevels(), aes(x= DATE_12, color = taxa)) + 
  stat_count( geom = "point", size= 4, alpha = 0.90, color = "purple") +
    ylab("Recordings per night") + xlab("Month") +
   ggtitle("Overall activity 2025") +
    scale_x_date(date_breaks = "1 month", , date_labels = "%b") +
    theme_minimal()


```

```{r enil 3}
#| echo: false
#| label: fig-enil-3
#| warning: false
#| fig-cap: Comparison of Enil activity in different sites
#| fig-subcap:
#| - "Activity on coast"
#| - "Activity around lakes"
#| - "Activity inland"
#| layout-ncol: 3
#| column: page


ggplot(cm %>% 
    filter(autoid == "EPTNIL",
            type == "coast") %>%
    droplevels(), aes(x= DATE_12, color = taxa)) + 
  stat_count( geom = "point", size= 4, alpha = 0.90, color = "purple") +
    ylab("Recordings per night") + xlab("Month") +
   ggtitle("Overall activity on the coast Enil") +
    scale_x_date(date_breaks = "1 month", , date_labels = "%b") +
    theme_minimal()

ggplot(cm %>% 
    filter(autoid == "EPTNIL",
            type == "lake") %>%
    droplevels(), aes(x= DATE_12, color = taxa)) + 
  stat_count( geom = "point", size= 4, alpha = 0.90, color = "purple") +
    ylab("Recordings per night") + xlab("Month") +
   ggtitle("Overall activity around lakes Enil") +
    scale_x_date(date_breaks = "1 month", , date_labels = "%b") +
    theme_minimal()
# 
# ggplot(cm %>% 
#     filter(autoid == "EPTNIL",
#             type == "inland") %>%
#     droplevels(), aes(x= DATE_12, color = taxa)) + 
#   stat_count( geom = "point", size= 4, alpha = 0.90, color = "purple") +
#     ylab("Recordings per night") + xlab("Month") +
#    ggtitle("Overall activity inland") +
#     scale_x_date(date_breaks = "1 month", , date_labels = "%b") +
#     theme_minimal()

```

## Nathusius' pipistrelle distribution

On the study area, some sites had clearly higher amounts of nathusius' pipistrelles. On the @fig-pnat-map below, we can see how many 5s clips of Auto-ID Pnat we had per site.

```{r pnat map}
#| echo: false
#| label: fig-pnat-map
#| warning: false
#| fig-cap: Number of Pnat calls per site

tilesURL <- "http://server.arcgisonline.com/ArcGIS/rest/services/Canvas/World_Light_Gray_Base/MapServer/tile/{z}/{y}/{x}"

basemap <- leaflet(width = "100%", height = "4000px") %>%
  addTiles(tilesURL)

basemap %>%
  addMinicharts(
    overview_summary$longitude, overview_summary$latitude,
    chartdata = overview_summary$pnatcalls, 
    showLabels = TRUE,
    width = 45
  )
```

The median date of activity did not differ that much based on the location, as is seen on @fig-pnat-ns below.

```{r pnat lat}


ggplot(filter(median_dates, Site != "CM-48")) +
  aes(x = `median`, y = `Site`) +
  geom_point(size = 3.35, 
             shape = "diamond") +
  scale_color_viridis_d(option = "cividis", direction = 1) +
  labs(x = "Median activity date", 
       y = "Site", title = "Median activity by site", subtitle = "From N to S") +
       theme_classic() +
  theme(legend.text = element_text(face = "bold"), legend.title = element_text(face = "bold"))

ggplot(cm_pnat_c) +
  aes(x = DATE_12) +
  geom_histogram(bins = 30L, fill = "#112446") +
  theme_minimal() +
  facet_wrap(vars(Site))


       
       
       
       
```

Diurnal activity patterns on the coast every half month from August to October.
```{r diur act pnat}
cm_pnat_c %>%
 filter(DATE_12 >= "2025-05-01" & DATE_12 <= "2025-07-31") %>%
 ggplot() +
  aes(x = HOUR_12) +
  geom_histogram(bins = 30L, fill = "#112446") +
  theme_minimal() +
  facet_wrap(vars(DATE))
cm_pnat_c %>%
  filter(DATE_12 >= "2025-05-01",
         DATE_12 <= "2025-07-30") %>%
  ggplot(aes(x = HOUR_12)) +
  geom_histogram(bins = 30, fill = "#112446") +
  theme_minimal() +
  facet_wrap(vars(HalfMonth))
```


```{r pnat lat map}
#| echo: false
#| eval: false
#| label: fig-pnat-map-ns
#| warning: false
#| fig-cap: Timing of Pnat calls per site

# data per day for recordings by site 
pnat_day <- cm_pnat %>%
  count(Site, DATE_12) %>%      # counts rows for each combination
  pivot_wider(
    names_from = DATE_12,
    values_from = n,
    values_fill = 0             # fill missing with 0
  )

# Reorder date columns
date_cols <- setdiff(names(pnat_day), "Site")  # All columns except site

# Convert column names to Date
ordered_cols <- date_cols[order(as.Date(date_cols))]

# Rebuild data frame with columns in date order
pnat_day <- pnat_day %>%
  select(Site, all_of(ordered_cols))

# add columns lat and long
pnat_day$lat <- overview_summary$latitude[match(pnat_day$Site, overview_summary$Site)]
pnat_day$long <- overview_summary$longitude[match(pnat_day$Site, overview_summary$Site)]

newdf <- pnat_day %>% select(-Site, -lat, -long)



basemap %>%
  addMinicharts(pnat_day$long, pnat_day$lat,
    chartdata= newdf,
    width = 45, height = 45
  )
