---
title: "Overview of data collected in the field season of 2025 for coastal monitoring"
author: "Maris Pärn"
format:
  html:
    embed-resources: true
editor: visual
---

```{r setup}
#| echo: false
#| warning: false

#setup

library(tidyverse)
library(ggplot2)
library(dplyr)
library(lubridate)
library(MetBrewer)
library(hms)
library(leaflet.minicharts)
library(leaflet)
library(tidyr)



# read necessary data in

overview_summary <- read.csv("overview_2025.csv")
cm <- read.csv("cm_2025.csv") 
missing_dates <- read.csv("missing_dates_2025.csv")
#missing_2024 <- read.csv("2024_missing_vetted.csv")

# overview figures

available_dates <- cm %>%
  mutate(ActualDate = as.Date(DATE_12)) %>%
  reframe(ActualDate = unique(ActualDate), .by = "Site")

complete_dates <- overview_summary %>% 
  mutate(
    BeginDate = as.Date(date_deployed),
    EndDate   = coalesce(as.Date(date_retrieved), as.Date("2025-10-22")) - 1,
    full_dates = map2(BeginDate, EndDate, ~ seq.Date(from = .x, to = .y, by = "day"))
  ) %>%
  select(Site, full_dates) %>%
  unnest(full_dates) %>%
  rename(ExpectedDate = full_dates)

```

## Overview

###Year in numbers

Number of detectors: 50\

Number of detectors that survived the season: 49\

Reasons of detector failure: seawater damage, storms, microphone losing sensitivity, teft.\

Date of first deployment: `r min(overview_summary$date_deployed)`\

Date of last deployment: `r max(overview_summary$date_deployed)`\

Date of first retrieval: `r min(overview_summary$date_retrieved)`\

Date of last retrieval: `r max(overview_summary$date_retrieved)`\

Days detectors were on the field: `r nrow(available_dates)`\

Days with missing data: `r nrow(missing_dates)`\

\% of missing data from all data: ca `r  100 * nrow(missing_dates)/ nrow(complete_dates)`\

Reasons of missing data:\

1\) No bats, no noise detected (later Nodata): 134\

2\) SD card full: 244\

3\) Batteries empty: 17\

4\) Detector dead: 113\

5\) Detector stolen: 55\

6\) SD card not formatted: 34\

Average time detector deployed: `r mean(overview_summary$days_active)`\

```{r preparing missing dates data}
#| echo: false
#| warning: false

complete_dates <- overview_summary |>  
  mutate(
    BeginDate = as.Date(date_deployed),
    EndDate   = as.Date(date_retrieved) - 1,
    full_dates = map2(BeginDate, EndDate, ~ seq.Date(from = .x, to = .y, by = "day"))
  ) |> 
  select(Site, full_dates) |> 
  unnest(full_dates) |> 
  rename(ExpectedDate = full_dates)

available_dates <- cm %>%
  mutate(ActualDate = as.Date(DATE_12)) %>%
  reframe(ActualDate = unique(ActualDate), .by = "Site")

available_dates <- available_dates %>%
  mutate(Type = "Observed")

missing_dates$X <- NULL

missing_dates <- missing_dates %>%
  rename(
    Date = ExpectedDate
  )

available_dates <- available_dates %>%
  rename(
    Date = ActualDate
  )

missing_dates <- missing_dates %>%
  mutate(Type = "Missing") %>% 
  mutate(Date=ymd(Date))


all_dates <- bind_rows(available_dates, missing_dates)
```

```{r data preparation for plots in this document}
#| echo: false
#| warning: false

# missing dates plot
  # add type NoData (aka no recordings) to Observed and Missing

missing_2025 <- missing_dates %>%
  mutate(ExpectedDate=mdy(ExpectedDate))

NoData <-  missing_2024 |> 
  filter(Battery.status == "OK", SD.card.status == "OK") |> 
  select(Site, ExpectedDate) |> 
  rename(
    Date = ExpectedDate
  ) |> 
  data.frame(Type = "NoData")

all_dates <- all_dates |> 
  left_join(NoData, by = c("Site", "Date"), suffix = c("", "_new")) |> 
  mutate(Type = if_else(!is.na(Type_new), Type_new, Type)) |> 
  select(-Type_new)

# noise plot

noise_coast <- cm[,c(13,16)]
noise_coast$type <- overview_summary$type[match(noise_coast$Site, overview_summary$Site)]

# autoid plot

auto_id_data <- cm[,c(10,13, 15, 16)]
auto_id_data$type <- overview_summary$type[match(auto_id_data$Site, overview_summary$Site)]

# combined bat calls plot

cm <- cm %>%
  mutate(DATE.12=ymd(DATE.12))

# activity plot
  # adding a column "type"

cm$type <- overview_summary$type[match(cm$Site, overview_summary$Site)]


```

```{r pnat data preparation}
#| echo: false
#| warning: false

# choose PIPNAT

cm_pnat <- cm %>%
  filter(autoid == "PIPNAT") 

cm_pnat_c <- cm_pnat |>
  filter(type == "coast") %>%
  filter(month(DATE.12) %in% c(8, 9, 10))

# add median dates

median_dates_all <- cm_pnat %>%
  group_by(Site) %>%
  summarise(median = median(DATE.12, na.rm = TRUE))

median_dates_all$latitude <-overview_summary$latitude[match(median_dates_all$Site, overview_summary$Site)]

median_dates_all <- median_dates_all %>%
  arrange(latitude) %>%
  mutate(Site = factor(Site, levels = Site))

cm_pnat_c <- cm_pnat_c %>%
  mutate(TIME=as_hms(TIME))

cm_pnat_c <- cm_pnat_c %>%
  mutate(TIME.12=as_hms(TIME.12))


# median(cm_pnat_c$DATE.12, na.rm = FALSE)

median_dates <- cm_pnat_c %>%
  group_by(Site) %>%
  summarise(median = median(DATE.12, na.rm = TRUE))

median_dates$latitude <-overview_summary$latitude[match(median_dates$Site, overview_summary$Site)]

median_dates <- median_dates %>%
  arrange(latitude) %>%
  mutate(Site = factor(Site, levels = Site))

cm_pnat_c <- cm_pnat_c %>%
  mutate(TIME=as_hms(TIME))

cm_pnat_c <- cm_pnat_c %>%
  mutate(TIME.12=as_hms(TIME.12))

# half a month

cm_pnat_c <- cm_pnat_c %>%
  mutate(
    HalfMonth = case_when(
      lubridate::day(DATE) <= 15 ~
        paste0(lubridate::year(DATE), "-", sprintf("%02d", lubridate::month(DATE)), "-H1"),
      TRUE ~
        paste0(lubridate::year(DATE), "-", sprintf("%02d", lubridate::month(DATE)), "-H2")
    )
  )

#check for duplicates
  #dim(cm_pnat_c[duplicated(cm_pnat_c$filename),])[1]

# 

```

## Deployment times

This is a graph that shows the days with missing data and how long the detectors were in the field.

```{r missing dates plot}
#| echo: false
#| label: fig-missing
#| warning: false
#| fig-cap: Missing dates 2024

ggplot(all_dates, aes(x = Date, y = Site, color = Type)) +
  geom_point(size = 2) +
  scale_color_manual(values = c("Observed" = "azure3", "NoData" = "darkgrey", "Missing" = "red")) +
  theme_minimal() +
  scale_y_discrete(expand = expansion(mult = c(0.01, 0.01)))+
  labs(title = "Observed and Missing Dates for 2024")


```
